---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
#install.packages("ggplot2")
library(ggplot2)

load(file = "../Data/Males_AllAges_0816.Rdata")
```

Set up your blank canvas:

```{r}
ggplot(data = CDC_Males_AllAges_0816_wide, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White))
```

Add the points:

```{r}
ggplot(data = CDC_Males_AllAges_0816_wide, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point()
```

Notice there are too many points. Use the `dplyr` function called `filter` to
subset to the firearm homicide and suicides only:

```{r}
install.packages("dplyr")
library(dplyr)
CDC_firearm_only <- CDC_Males_AllAges_0816_wide %>% filter(type == "Firearm") 

ggplot(data = CDC_firearm_only, aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point()
```

Link color to state gun prevalence. Try both continuous and categorical variables
to see the difference. RememberL you need to do this inside the `aes()` function!
Try putting it outside the aes() and see what happens. 

```{r}
#categorical version
#why is categorical data colours in a non ordered way?
#do we want it to be ordered here?
ggplot(data = CDC_firearm_only, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(col = gun.house.prev.category))

# continuous version
# what is misleading here about the default?
# the darkest colours correspond to the lower values. Confusing!
ggplot(data = CDC_firearm_only, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(col = gun.house.prev.category)) 
```

Set the colors manually. Do this inside of the scale function:

```{r}
ggplot(data = CDC_firearm_only, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(col = gun.house.prev.category)) +
  scale_color_manual(values = c('#fef0d9','#fdcc8a','#fc8d59','#d7301f'))
```

Link to size:

```{r}
ggplot(data = CDC_firearm_only, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(col = gun.house.prev.category, size = average.pop.white)) +
  scale_color_manual(values = c('#fef0d9','#fdcc8a','#fc8d59','#d7301f')) 
```

Tell the legend for size where to show the breaks:

```{r}
ggplot(data = CDC_firearm_only, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(col = gun.house.prev.category, size = average.pop.white)) +
  scale_color_manual(values = c('#fef0d9','#fdcc8a','#fc8d59','#d7301f')) +
  scale_size_area(breaks = c(200000, 500000, 1500000, 3000000, 7000000))
```

Make the max size of the circles larger:

```{r}
ggplot(data = CDC_firearm_only, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(col = gun.house.prev.category, size = average.pop.white)) +
  scale_color_manual(values = c('#fef0d9','#fdcc8a','#fc8d59','#d7301f')) +
  scale_size_area(breaks = c(200000, 500000, 1500000, 3000000, 7000000), 
                  max_size = 15)
```

Lots of changes in this next plot:

* add the x and y axis labels inside `labs()`
* change the type of plotting point using pch. Then need to use fill instead of color for pch=21 (since this pch has both a fill and an outline)

```{r}
ggplot(data = CDC_firearm_only, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(fill = gun.house.prev.category, size = average.pop.white), pch = 21) +
  scale_fill_manual(values = c('#fef0d9','#fdcc8a','#fc8d59','#d7301f')) +
  scale_size_area(breaks = c(200000, 500000, 1500000, 3000000, 7000000), 
                  max_size = 15) + 
  theme_bw() +
  labs(x = "White suicide rate (per 100,000 per year)", 
       y = "White homicide rate (per 100,000 per year)")
```

Add state labels with geom_text() or geom_label(). Try both and see how they differ.

```{r}
ggplot(data = CDC_firearm_only, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(fill = gun.house.prev.category, size = average.pop.white), pch = 21) +
  scale_fill_manual(values = c('#fef0d9','#fdcc8a','#fc8d59','#d7301f')) +
  scale_size_area(breaks = c(200000, 500000, 1500000, 3000000, 7000000), 
                  max_size = 15) + 
  theme_bw() +
  labs(x = "White suicide rate (per 100,000 per year)", 
       y = "White homicide rate (per 100,000 per year)") +
  geom_label(aes(label = ST))
```

Use the package ggrepel to repel these labels away from one another and away from
the data so they don't appear so crowded. Need to change geom_text (or geom_label)
to geom_text_repel

```{r}
library(ggrepel)

ggplot(data = CDC_firearm_only, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(fill = gun.house.prev.category, size = average.pop.white), pch = 21) +
  scale_fill_manual(values = c('#fef0d9','#fdcc8a','#fc8d59','#d7301f')) +
  scale_size_area(breaks = c(200000, 500000, 1500000, 3000000, 7000000), 
                  max_size = 15) + 
  theme_bw() +
  labs(x = "White suicide rate (per 100,000 per year)", 
       y = "White homicide rate (per 100,000 per year)") +
  geom_text_repel(aes(label = ST))
```

Need to add the spearman's rank to the plot. Use this `dplyr` code to calculate
the spearman's rank statistic and call it rho

```{r}
corr <- cor.test(x = CDC_firearm_only %>% filter(! ST %in% c("US", "HI", "DC")) %>% select(adjusted.homicide.White) %>% unlist(), 
                 y = CDC_firearm_only %>% filter(! ST %in% c("US", "HI", "DC")) %>% select(adjusted.suicide.White) %>% unlist(), 
                 method = 'spearman')

corr

rho <- corr$estimate
```

Glue is a great package for glueing together words with variables:

```{r}
library(glue)
glue("The Spearman's rank coefficient is:{rho}")

glue("The Spearman's rank coefficient is:{round(rho, 2)}")
```

Add rho to the plot using geom_text:
You need to supply x and y in this aes() to tell the text where to plot it
You need to also say check_overlap = T or else it will plot it for each row of the data and appear bolded (try removing check_overlap =  T)

```{r}
ggplot(data = CDC_firearm_only, 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(fill = gun.house.prev.category, size = average.pop.white), pch = 21) +
  scale_fill_manual(values = c('#fef0d9','#fdcc8a','#fc8d59','#d7301f')) +
  scale_size_area(breaks = c(200000, 500000, 1500000, 3000000, 7000000), 
                  max_size = 15) + 
  theme_bw() +
  labs(x = "White suicide rate (per 100,000 per year)", 
       y = "White homicide rate (per 100,000 per year)") +
  geom_text_repel(aes(label = ST)) +
  geom_text(aes(x = 25, y = 0.75, label = glue("Spearman's rho: {round(rho, 2)}")), check_overlap = T)
```

Make the legend pretty. See slides explaining what each new bit of code does:

```{r}
ggplot(data = CDC_firearm_only %>% filter(!ST %in% c("US", "HI", "DC")), 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(fill = gun.house.prev.category, size = average.pop.white), pch = 21) +
  scale_fill_manual(values = c('#fef0d9','#fdcc8a','#fc8d59','#d7301f')) +
  scale_size_area(breaks = c(200000, 500000, 1500000, 3000000, 7000000), 
                  labels = c("200k", "500k", "1.5m", "3m", "7m"),
                  max_size = 15) + 
  theme_bw() +
  labs(x = "White suicide rate (per 100,000 per year)", 
       y = "White homicide rate (per 100,000 per year)") +
  geom_text_repel(aes(label = ST)) +
  geom_text(aes(x = 25, y = 0.75, label = glue("Spearman's rho: {round(rho, 2)}")), check_overlap = T) +
  guides(fill = guide_legend(title = "Gun ownership", override.aes = list(size = 5), order = 1),
         size = guide_legend(title = "White population"), order = 2)
  

```

Don't forget to save your data! This code will only work if you add a Plots folder 
inside of your main folder!

```{r save-figures, echo=F}
figure.3 <- ggplot(data = CDC_firearm_only %>% filter(!ST %in% c("US", "HI", "DC")), 
       aes(x = adjusted.suicide.White, y = adjusted.homicide.White)) +
  geom_point(aes(fill = gun.house.prev.category, size = average.pop.white), pch = 21) +
  scale_fill_manual(values = c('#fef0d9','#fdcc8a','#fc8d59','#d7301f')) +
  scale_size_area(breaks = c(200000, 500000, 1500000, 3000000, 7000000), 
                  labels = c("200k", "500k", "1.5m", "3m", "7m"),
                  max_size = 15) + 
  theme_bw() +
  labs(x = "White suicide rate (per 100,000 per year)", 
       y = "White homicide rate (per 100,000 per year)") +
  geom_text_repel(aes(label = ST)) +
  geom_text(aes(x = 25, y = 0.75, label = glue("Spearman's rho: {round(rho, 2)}")), check_overlap = T) +
  guides(fill = guide_legend(title = "Gun ownership", override.aes = list(size = 5), order = 1),
         size = guide_legend(title = "White population"), order = 2)

ggsave(filename = "../Plots/figure3.jpeg", plot = figure.3, device = "jpeg", width = 8, height = 6, units = "in")
```